---
title: "Regression Analysis"
output: html_document
---

<br>

```{r include=FALSE, message=FALSE}
library(tidyverse)
library(performance)
library(modelr)
library(mgcv)
library(corrplot)
library(leaps)
library(glmnet)
library(rsample)
library(caret)
library(patchwork)
library(Metrics)
library(MASS)
library(glmnet)

knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 6,
  fig.asp = .7,
  dpi = 800,
  out.width = "80%"
)

theme_set(theme_minimal())

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r include=FALSE, message=FALSE}
le_df = 
  read_csv("./data/Dummy_Data.CSV")
```

<br>

## Correlation Matrix

```{r}
cm_df =
  le_df %>%
  dplyr::select(-c(country,continent,year))

corrplot(cor(cm_df), type = "upper", diag = FALSE, method = "square", addCoef.col = "black", number.cex = .3, tl.col = "black", tl.cex = .7)
```

<br>

<hr>

## Model building 

### Model Based on Prior Research

```{r}
mod_1 = lm(le ~ status_Developed + percent_exp + HDI + adult_mort + infantdeaths + schooling + gdp + hiv_aids + gdp * percent_exp, data = le_df)
summary(mod_1)
```


```{r}
check_collinearity(mod_1)
```


<br>
<br>

### Models Based on Automated Model Selection

```{r include = FALSE}
full_model = lm(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + measles + bmi + under_five + polio_high + total_exp + diphtheria_high  + hiv_aids + gdp + population + thin_1_19 + thin_5_9 + HDI + schooling, data = le_df)
```


**Stepwise Regression:**

```{r}
step_model = stepAIC(full_model, direction = "both", trace = FALSE)
summary(step_model)
check_collinearity(step_model)
```

*Final Stepwise Model*
```{r}
## remove insignificant and high correlation variables (under_five, infantdeaths)
mod_2 = lm(le ~ status_Developed + adult_mort + bmi + polio_high + diphtheria_high + hiv_aids + gdp + thin_1_19 + HDI + schooling, data = le_df)
```


<br>
<br>

**Forward/Backward Selection:**

*Backward Elimination*
```{r}
back_mod = stepAIC(full_model, direction = "backward", 
                      trace = FALSE, scope = formula(full_model))
summary(back_mod)
check_collinearity(back_mod)

backward = lm(le ~ status_Developed + adult_mort + bmi + polio_high + diphtheria_high + hiv_aids + gdp + thin_1_19 + HDI + schooling, data = le_df)
```

*Forward Selection*
```{r}
forward_mod = stepAIC(full_model, direction = "forward", trace = FALSE, scope = formula(full_model))
summary(forward_mod)
check_collinearity(forward_mod)

forward = lm(le ~ status_Developed + adult_mort + infantdeaths + bmi + polio_high + diphtheria_low + hiv_aids + gdp + HDI + schooling, data = le_df)
```

<br>
<br>


**Best Subsets Regression:**

<br>

*Best Fitting Models Based on the Numbers of Predictors*
```{r include=FALSE}
sub_fit = regsubsets(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + measles + bmi + under_five + polio_high + total_exp + diphtheria_high  + hiv_aids + gdp + population + thin_1_19 + thin_5_9 + HDI + schooling, data = le_df)

best_summary = summary(sub_fit)
best_summary_df = 
  as.data.frame(best_summary$outmat) %>% 
  remove_rownames() %>% 
  add_column(" " = c(1:8), .before = "status_Developed")

table = as.data.frame(t(best_summary_df[-1])) %>% 
  rename("1" = "V1", "2" = "V2", "3" = "V3","4" = "V4", "5" = "V5", "6" = "V6","7" = "V7", "8" = "V8")
```

```{r}
table %>% 
     knitr::kable()
```

<br>

*Comparison Based on Model Selection Criteria*
```{r}
Model_comparison =
  tibble(
	n_pred = c(1:8),
	"Adjusted R2" = best_summary$adjr2,
	"Adj R2" = best_summary$adjr2,
	"BIC" = best_summary$bic,
	"Cp" = best_summary$cp,
	"RSE" = best_summary$rsq,
	"RSS" = best_summary$rss
	) %>%
	gather(key = "Statistics", value = "value", 2:6) %>%
	ggplot(aes(x = n_pred, y = value)) +
	  geom_point() +
	  geom_line() +
	  facet_grid(Statistics ~ ., scales = "free_y") +
	  labs(
	  	x = "Number of Predictors",
	  	y = "Values",
	  	title = "Model Selection Criteria Comparison"
	  )

Model_comparison
```

*Final Best Subsets Model*
```{r}
mod_3 = lm(le ~ adult_mort + under_five + polio_high + hiv_aids + gdp + HDI + schooling, data = le_df)
```

<br>
<br>

**Ridge Regression:**
```{r}
#define response variable
y = le_df$le

#define matrix of predictor variables
x = data.matrix(le_df[, c('status_Developed', 'adult_mort', 'infantdeaths', 'alcohol', 'percent_exp', 'hepatitis_b_high', 'measles', 'bmi', 'under_five', 'polio_high', 'total_exp', 'diphtheria_high', 'hiv_aids', 'gdp', 'population', 'thin_1_19', 'thin_5_9', 'HDI', 'schooling')])

#fit ridge regression model
ridge_model = glmnet(x, y, alpha = 0)
  
#perform k-fold cross-validation to find optimal lambda value
cv_ridge_model = cv.glmnet(x, y, alpha = 0)

best_lambda = cv_ridge_model$lambda.min
best_lambda

best_ridge_model = glmnet(x, y, alpha = 0, lambda = best_lambda)
coef(best_ridge_model)

#produce plot of test MSE by lambda value
plot(cv_ridge_model) 
```


<br>
<br>


**Lasso Regression:**

```{r}
#define response variable
y = le_df$le

#define matrix of predictor variables
x = data.matrix(le_df[, c('status_Developed', 'adult_mort', 'infantdeaths', 'alcohol', 'percent_exp', 'hepatitis_b_high', 'measles', 'bmi', 'under_five', 'polio_high', 'total_exp', 'diphtheria_high', 'hiv_aids', 'gdp', 'population', 'thin_1_19', 'thin_5_9', 'HDI', 'schooling')])

#perform k-fold cross-validation to find optimal lambda value
cv_lasso_model = cv.glmnet(x, y, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lasso_lambda = cv_lasso_model$lambda.min
best_lasso_lambda

best_lasso_model = glmnet(x, y, alpha = 1, lambda = best_lasso_lambda)
coef(best_lasso_model)

new_x = data.matrix(le_df[, c('status_Developed', 'adult_mort', 'infantdeaths', 'alcohol', 'percent_exp', 'hepatitis_b_high', 'bmi', 'under_five', 'polio_high', 'total_exp', 'diphtheria_high', 'hiv_aids', 'gdp', 'thin_1_19', 'thin_5_9', 'HDI', 'schooling')])

plot(cv_lasso_model)
```

<br>
<br>
<hr>

## Cross Validation

```{r include=FALSE}
set.seed(2)

cv_df = 
  crossv_mc(le_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>% 
  mutate(
    mod_1 = map(train, ~lm(formula = formula(mod_1), data = .x)),
    mod_2 = map(train, ~lm(formula = formula(mod_2), data = .x)),
    mod_3 = map(train, ~lm(formula = formula(mod_3), data = .x)),
    
    mod_4 = map(train, ~train(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + measles + bmi + under_five + polio_high + total_exp + diphtheria_high + hiv_aids + gdp + population + thin_1_19 + thin_5_9 + HDI + schooling, data = training(initial_split(le_df, prop = 0.75))[,-1], trControl=trainControl(method="cv", number = 5), method="glmnet", preProcess = c("center", "scale"),tuneGrid = expand.grid(alpha = 0, lambda = best_lambda)), data = .x),
    
    mod_5 = map(train, ~train(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + bmi + under_five + polio_high + total_exp + diphtheria_high + hiv_aids + gdp + thin_1_19 + thin_5_9 + HDI + schooling, data = training(initial_split(le_df, prop = 0.75))[,-1], trControl=trainControl(method="cv", number = 5), method="glmnet", preProcess = c("center", "scale"),tuneGrid = expand.grid(alpha = 1, lambda = best_lasso_lambda)), data = .x)) %>% 
  
  mutate(
    rmse_1 = map2_dbl(mod_1, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_2 = map2_dbl(mod_2, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_3 = map2_dbl(mod_3, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_4 = map2_dbl(mod_4, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_5 = map2_dbl(mod_5, test, ~modelr::rmse(model = .x, data = .y))
    ) %>% 
  mutate(
    mae_1 = map2_dbl(mod_1, test, ~modelr::mae(model = .x, data = .y)),
    mae_2 = map2_dbl(mod_2, test, ~modelr::mae(model = .x, data = .y)),
    mae_3 = map2_dbl(mod_3, test, ~modelr::mae(model = .x, data = .y)),
    mae_4 = map2_dbl(mod_4, test, ~modelr::mae(model = .x, data = .y)),
    mae_5 = map2_dbl(mod_5, test, ~modelr::mae(model = .x, data = .y))
    ) %>% 
  mutate(
    rsq_1 = map2_dbl(mod_1, test, ~modelr::rsquare(model = .x, data = .y)),
    rsq_2 = map2_dbl(mod_2, test, ~modelr::rsquare(model = .x, data = .y)),
    rsq_3 = map2_dbl(mod_3, test, ~modelr::rsquare(model = .x, data = .y)),
    rsq_4 = map2_dbl(mod_4, test, ~modelr::rsquare(model = .x, data = .y)),
    rsq_5 = map2_dbl(mod_5, test, ~modelr::rsquare(model = .x, data = .y))
    )
```


### Cross-validating with Violin Plots
```{r include=FALSE}
rmse_table = 
  cv_df %>%
  pivot_longer(
    rmse_1:rmse_5,
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model),
         model = fct_recode(model, "Model 1" = "1",
                                   "Model 2" = "2",
                                   "Model 3" = "3",
                                   "Model 4" = "4",
                                   "Model 5" = "5" ))
rsq_table = 
  cv_df %>% 
  pivot_longer(
    rsq_1:rsq_5,
    names_to = "model", 
    values_to = "rsq",
    names_prefix = "rsq_") %>% 
  mutate(model = fct_inorder(model),
         model = fct_recode(model, "Model 1" = "1",
                                   "Model 2" = "2",
                                   "Model 3" = "3",
                                   "Model 4" = "4",
                                   "Model 5" = "5" ))
mae_table = 
  cv_df %>% 
  pivot_longer(
    mae_1:mae_5,
    names_to = "model", 
    values_to = "mae",
    names_prefix = "mae_") %>% 
  mutate(model = fct_inorder(model),
         model = fct_recode(model, "Model 1" = "1",
                                   "Model 2" = "2",
                                   "Model 3" = "3",
                                   "Model 4" = "4",
                                   "Model 5" = "5" ))
```


```{r}
rmse_plot =
  rmse_table %>% 
  ggplot(aes(x = model, y = rmse, fill = model)) + 
  geom_violin() + 
  labs(x = "",
       y = "RMSE") +
  scale_fill_brewer(type = 'seq', palette = 'Set2') +
  theme(legend.position = 'none',
        plot.title = element_text(hjust = 0.5))

rsq_plot =
  rsq_table %>% 
  ggplot(aes(x = model, y = rsq, fill = model)) + 
  geom_violin() + 
  labs(x = "",
       y = "RSquared") +
  scale_fill_brewer(type = 'seq', palette = 'Set2') +
  theme(legend.position = 'none',
        plot.title = element_text(hjust = 0.5))

mae_plot =
  mae_table %>% 
  ggplot(aes(x = model, y = mae, fill = model)) + 
  geom_violin() + 
  labs(x = "",
       y = "MAE") +
  scale_fill_brewer(type = 'seq', palette = 'Set2') +
  theme(legend.position = 'none',
        plot.title = element_text(hjust = 0.5))

rmse_plot/rsq_plot/mae_plot
```

<br>
<br>

### Cross Validation Table
```{r message=FALSE}
mean_rmse =
  rmse_table %>% 
  group_by(model) %>% 
  summarize(mean_rmse = mean(rmse))

mean_mae = 
  mae_table %>%
  group_by(model) %>% 
  summarize(mean_mae = mean(mae))

mean_rsquare = 
  rsq_table %>% 
  group_by(model) %>% 
  summarize(mean_rsq = mean(rsq))

Model_comparison_table =
  full_join(inner_join(mean_rmse, mean_rsquare), mean_mae) %>% 
  knitr::kable()

Model_comparison_table 
```


<br>
<hr>

## Results

```{r}
mod_1 %>% 
  broom::tidy() %>% 
  knitr::kable()
```

