---
title: "Hypothesis Test"
output: html_document
---

<style type="text/css">

h1.title {
  font-size: 50px;
  text-align: left;
  font-family: Papyrus;
  font-weight: bold;
  font-variant: small-caps
}
</style>

<br>

```{r include=FALSE, message=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(corrplot)
library(ggpubr)
library(performance) 

knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 6,
  fig.asp = .7,
  dpi = 800,
  out.width = "80%"
)

theme_set(theme_minimal())

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r include=FALSE, message=FALSE}
le_df = 
  read_csv("./data/Complete_Data.CSV") %>% 
  rename("HDI" = "incomecompositionofresources", "percent_exp" = "percentageexpenditure", "total_exp" = "totalexpenditure", "le" = "lifeexpectancy", "adult_mort" = "adult_mortality", "thin_1_19" = "thinness1_19years", "thin_5_9" = "thinness5_9years", "under_five" = "under_fivedeaths")
```

There are countless factors that affects country's average life expectancy and how long an individual will be expected to live. In continuation of the exploratory analysis, we want to dive more into the dataset to further investigate the variables and possibly excavate some interesting findings that we don’t normally aware of. 

<br>

### **Development Status**

When exploring the Development status variable, the one-way analysis of variance (ANOVA) is used in order to have an understanding on how the two different groups (Developed and Developing) respond. And from the result we obtained, a conclusion was drawn that there is statistically significant differences between the means of Developed countries and Developing countries. 

Subsequently, we want to see if the dataset provides envidence for Developed countries having greater mean life expectancy than the Developing countries. 

<center>
$H_{0}: \mu_{Developed} = \mu_{Developing}$ \
$H_{1}: \mu_{Developed} > \mu_{Developing}$ 
</center>
<br>

We will perform two-sample T-Test since the true variance of the populations is unknown. And also use F-test to test if the variances of the two populations are equal. 

```{r,include=FALSE}
## mean life expectancy of each country from 2000 - 2015
Developed = 
  le_df %>% 
  filter(status == "Developed") %>% 
  group_by(country) %>% 
  summarize(Mean_LE = mean(le))

Developing = 
  le_df %>% 
  filter(status == "Developing") %>% 
  group_by(country) %>% 
  summarize(Mean_LE = mean(le))
```

```{r, message=FALSE, fig.width=5}
var.test(Developed$Mean_LE, Developing$Mean_LE) %>% 
  broom::tidy() %>% 
  select(c(4,5,8,9)) %>% 
  knitr::kable(table.attr = "style='width:30%;'")
```

* We obtain a p-value of $2.21e^{-8}$ which is less than $\alpha = 0.05$. Hence we reject the null hypothesis and conclude that the variance of two populations is not equal.

```{r}
t.test(Developed$Mean_LE, Developing$Mean_LE, alternative = "greater", var.equal = FALSE) %>% 
  broom::tidy() %>% 
  select(c(4,5,6,9)) %>% 
  knitr::kable(table.attr = "style='width:30%;'")
```

* At significance level of 0.05, we reject the $H_{0}$ since $p-value < 2.2e^{-16} < \alpha = 0.05$. Thus, we conclude that the life expectancy for developed countries is greater than the developing countries.

<br>

### **Income composition of resources (HDI)**

The Human Development Index (HDI) is a statistic composite index of average achievement in key dimensions of human development: health (life expectancy),  education (knowledge), and per capita income indicators (standard of living); it measures how good a country is at using its resources. By definition, life expectancy is one of the aspects of HDI measure, we want to see if there's strong relationship between HDI and life expectancy; if countries that spend a higher proportion of their resources on human development would have a higher life expectancy.

```{r include=FALSE}
human_develop = 
  le_df %>% 
  group_by(country) %>% 
  summarize(mean_LE = mean(le), 
            mean_HDI = mean(HDI))
```

```{r, message=FALSE}
human_develop %>% 
  ggplot(aes(x = mean_HDI, y = mean_LE)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(aes(label = ..r.label..), method = "pearson",  label.y = 75,color = "red") +
  labs(title = "Mean Life Expectancy VS. Mean HDI",
       x = "Mean HDI",
       y = "Mean Life Expectancy (Years)")
```

* We obtain a correlation coefficient value of r = 0.79, which shows a significant positive correlation between life expectancy and HDI. This implies that countries with higher income composition of resources for human development and utilizes resources effectively are more likely to have higher life expectancy. Thus, countries could spend more on human development in order to achieve higher life expectancy. 

<br>

### **Infant deaths and Under-five deaths**

From the [correlation matrix](regression_files/figure-html/unnamed-chunk-3-1.png), the variable infant deaths and under-five deaths are highly correlated. Hence, we want to compare the proportions of the two variables (infant deaths and under-five deaths) using two-proportions z-test.

<center>
$H_{0}: p_{infant} = p_{under five}$ \
$H_{1}: p_{infant} \ne p_{under five}$
</center>

```{r, include=FALSE}
infant_underfive_deaths = 
  le_df %>% 
  group_by(country) %>% 
  summarize(mean_LE = mean(le),
            mean_infant = mean(infantdeaths),
            mean_underfive = mean(under_five))

avg_infants = mean(infant_underfive_deaths$mean_infant)
avg_under = mean(infant_underfive_deaths$mean_underfive)
x = c(avg_infants,avg_under)
n = c(1000,1000)
```

```{r,message=FALSE}
prop.test(x, n, correct = FALSE) %>% 
  broom::tidy() %>% 
  select(c(3,4,5,8)) %>% 
  knitr::kable(digits = 4)
```

* At 0.05 significance level, we have p-value = 0.1709 $> \alpha = 0.05$, thus, we fail to reject the null hypothesis and there is insufficient evidence to conclude that there is significant difference in the two independent proportions. Hence, we need to take into consideration of possible collinearity between the infant deaths variable and the under-five deaths variable when building the regression model.

* We can see from the [correlation matrix](regression_files/figure-html/unnamed-chunk-3-1.png) that the correlation coefficient between infant deaths and life expectancy is negative (r = -0.2), and similarly between under-five deaths and life expectancy (r = -0.22), which means that when infant deaths or under-five deaths increase, life expectancy will eventually decreases.

<br>

### **Alcohol**

A recent cohort study showed that light-to-moderate alcohol intake is positively associated with longevity [(Brandts et al., 2020)](https://academic.oup.com/ageing/article/49/3/395/5730334). Another study conducted by a group of Harvard researchers on how lifestyle factors affect life expectancy showed evidence that people who take a few alcoholic drinks a week tends to live more years free of diabetes and heart disease than those who don't drink or binge drink [(Li et al., 2018)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6207481/).

We want to see if life expectancy shows any strong correlation with drinking alcohol using our data.

```{r, include=FALSE}
alc = 
  le_df %>% 
  group_by(country) %>% 
  summarize(mean_LE = mean(le), 
            mean_alc = mean(alcohol))
```

```{r, message=FALSE}
alc %>% 
  ggplot(aes(x = mean_alc, y = mean_LE)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(aes(label = ..r.label..), method = "pearson", label.x = 10, label.y = 60, color = "red") +
  labs(title = "Mean Life Expectancy VS. Mean Alcohol Consumption",
       x = "Alcohol Consumption (per capita (15+) consumption (in litres of pure alcohol))",
       y = "Life Expectancy (Years)") 
```

* Surprisingly, in accordance with the studies, the scatter plot demonstrates that alcohol consumption has a positive correlation (r = 0.44) with life expectancy.

```{r, include=FALSE}
alc_mortality = 
  le_df %>% 
  group_by(country) %>% 
  summarize(mean_mortality = mean(adult_mort), 
            mean_alc = mean(alcohol))
```

```{r, message=FALSE}
alc_mortality %>% 
  ggplot(aes(x = mean_alc, y = mean_mortality)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method = "pearson", label.x = 8, color = "red") +
  labs(title = "Mean Adult Mortality Rates VS. Mean Alcohol Consumption",
       x = "Alcohol Consumption (per capita (15+) consumption (in litres of pure alcohol))",
       y = "Adult Mortality Rates (per 1000 population)") 
```

* Although there's no strong correlation between adult mortality rates and alcohol consumption, but having a p-value = 0.00083 we see that the correlation coefficient would still be significant. Even though, we can't reach a strong conclusion regarding the correlation, interestingly, there's evidence showing that alcohol consumption is positively associated (r = 0.44) with life expectancy and negatively associated (r = -0.24) with adult mortality. 

<br>

### **Correlation between immunization and other health related factors**

By understanding correlations between variables, we could gain knowledge on the factors that will affect life expectancy the most.

```{r}
cor_df =
  le_df %>%
  dplyr::select(c(le, hepatitis_b, measles, polio, diphtheria, hiv_aids, thin_1_19, thin_5_9))

corrplot(cor(cor_df), type = "upper", diag = FALSE, method = "square", addCoef.col = "black", number.cex = 1, tl.col = "black", tl.cex = .8)
```

* Life Expectancy has moderate negative correlation with HIV/AIDS, Thinness_1_19_years, and Thinness_5_9_years. And has low negative correlation with Measles.

* Life Expectancy moderate positive correlation with Diphtheria, Polio and Hepatitis_B

<br>

In the US, nearly all Americans have healthcare or health insurance, the accessibility to health care undoubtedly decreases mortality and increases life expectancy. A study conducted for East African countries have shown that there is a strong positive association between total healthcare expenditures and life expectancy; also, a negative relationship between healthcare expenditures and number of infant and under-five deaths [Bein et al., 2017](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5636241/).


```{r, include=FALSE}
toalexp = 
  le_df %>% 
  group_by(country) %>% 
  summarize(mean_LE = mean(le), 
            mean_toalexp = mean(total_exp))

exp_inf = 
  le_df %>% 
  group_by(country) %>% 
  summarize(mean_inf = mean(infantdeaths), 
            mean_toalexp = mean(total_exp))

exp_underfive = 
  le_df %>% 
  group_by(country) %>% 
  summarize(mean_under = mean(under_five), 
            mean_toalexp = mean(total_exp))
```

```{r, message=FALSE, fig.height=8, fig.width=16}
toalexp %>% 
  ggplot(aes(x = mean_toalexp, y = mean_LE)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method = "pearson",label.y = 85, color = "red") +
  stat_regline_equation(label.y = 82, aes(label = ..rr.label..), color = "red") +
  labs(title = "Mean Healthcare Expenditures VS. Mean Life Expectancy",
       x = "Healthcare Expedicture (% of Total Expenditure)",
       y = "Life Expectancy (Years)") 

B = exp_inf %>% 
  ggplot(aes(x = mean_toalexp, y = mean_inf)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(aes(label = ..r.label..), method = "pearson", color = "red") +
  labs(title = "Mean Healthcare Expenditures VS. Mean Infant Deaths",
       x = "Healthcare Expedicture (% of Total Expenditure)",
       y = "Infant Deaths") 

C = exp_underfive %>% 
  ggplot(aes(x = mean_toalexp, y = mean_under)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(aes(label = ..r.label..), method = "pearson", color = "red") +
  labs(title = "Mean Healthcare Expenditures VS. Mean Under-five Deaths",
       x = "Healthcare Expedicture (% of Total Expenditure)",
       y = "Under-five Deaths") 

gridExtra::grid.arrange(B, C, nrow = 1)
```

* Our dataset also shows evidence that expenditure spend on health is positively associated (r = 0.22) with life expectancy and negatively associated with infant (r = -0.16) and under-five (r = -0.16) deaths.

Now, it will be interesting to see if countries with low life expectancy (<69 (mean)) spend less of their government expenditure on health and if those countries should increase their total expenditure on healthcare to increase life expectancy.

```{r}
health_Low = 
  le_df %>% 
  filter(le < 69) %>% 
  group_by(country) %>% 
  summarize(Mean_exp = mean(total_exp))

health_High = 
  le_df %>% 
  filter(!le < 69) %>% 
  group_by(country) %>% 
  summarize(Mean_exp = mean(total_exp))

t.test(health_Low$Mean_exp, health_High$Mean_exp) %>% 
  broom::tidy() %>% 
  select(c(4,5,6,9)) %>% 
  knitr::kable(table.attr = "style='width:30%;'")
```

* The p-value we get from the t-test is 0.047 which is less than significance level of 0.05, thus, we will reject the null hypothesis and conclude that countries with high life expectancy spend higher proportion of their government expenditure on healthcare than countries with low life expectancy. 

* Also, from the plot above, we see that total expenditure on health as a percentage of of total government expenditure is a factor contribute in predicting life expectancy since we have a p-value = 0.0025. However, we can infer that solely total healthcare expenditure factor will not accurately predict life expectancy since we obtain a pretty low $R^{2}$ value of 0.047. Therefore, we want to look at other factors along with total expenditures and see how they are going to affect the life expectancy. 

<br>

### **Correlation between socioeconomic and behavioral factors**

Many demographic, socioeconomic, and behavioral factors can be linked with individual's life expectancy. In the United States, higher income was associated with longer life expectancy, and there are increasing differences in life expectancy across income groups [Chetty et al., 2016](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4866586/). We will conduct linear regression analysis based on SES and behavioral factors such as `alcohol`, `percent_exp`, `total_exp`, `bmi`, `population`, `HDI`, `schooling`, and `GDP` to have a deeper understanding on how these factors influence life expectancy of a country.  

```{r}
SES = lm(le ~ alcohol + percent_exp + total_exp + bmi + population + HDI + schooling, data = le_df)

summary(SES)
vif1 = check_collinearity(SES)
```

* We see that `total_exp` and `population` are both insignificant. 

```{r}
add_GDP = lm(le ~ alcohol + percent_exp + total_exp + bmi + population + HDI + schooling + gdp, data = le_df)
vif2 = check_collinearity(add_GDP) 
summary(add_GDP)
```


* Surprisingly, `percent_exp` becomes insignificant along with `total_exp` and `population` variables when `gdp` is added to the regression model. There's possibility that some of the variables are strongly correlated (`percent_exp` and `gdp`).

```{r}
ses_beh_df =
  le_df %>%
  dplyr::select(c(le, alcohol, percent_exp, total_exp, bmi,gdp, HDI, population, schooling))

corrplot(cor(ses_beh_df), type = "upper", diag = FALSE, method = "square", addCoef.col = "black", number.cex = 0.5, tl.col = "black", tl.cex = .8)
```

* `percent_exp` and `gdp` are strongly correlated as well, which explains why `percent_exp` becomes insignificant when `gdp` factor was added to the model. Income composition of resources (`HDI`) is highly correlated with `life expectancy` and `schooling`, it shows evidence that it's a composite variable. 

```{r}
remove_HDI = lm(le ~ alcohol + percent_exp + total_exp + bmi + population + schooling + gdp, data = le_df)

summary(remove_HDI)

vif3 = check_collinearity(remove_HDI) ## all vif drop below 2
```

* VIF all fell below 2, HDI could be a combination of other variables.  

